---
title: "Driver offences NZ"
output:
  pdf_document: default
  html_notebook: default
---

#####Driving offence data for NZ - sourced from police.govt.nz

Original file -

The following code reads different sheets from the spreadsheet, combining data for offences and fines by each offence type. There's a main data frame `offences_and_fines` which contains this set.

Note - Sheets are hardcoded, hence tests must be done to double check total values.


```{r load_libraries, include=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library("readxl"))
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library("ggthemes"))
suppressPackageStartupMessages(library("lubridate"))
```


```{r globals, echo=TRUE}

# General pattern for offence columns, the gap is to exclude year for consistent renaming. Update the range as new quarterly data comes through
select_columns <- c(1:14,16:27,29:40,42:53,55:66,68:79,81:92,94:102)

column_names <- c("area","sub_area",as.character(seq(as.Date("2009/1/1"), as.Date("2016/9/1"), "months")))

# General function to restructure and reassign frames from different sheets
restructureData <- function(dataFrame, typeOfOffence, countCategory){
  column_names <- c("area","sub_area",as.character(seq(as.Date("2009/1/1"), as.Date("2016/9/1"), "months")))
  
  names(dataFrame) <- column_names
  
  dataFrame <- dataFrame %>% 
    gather('month_year', 'count', -area, -sub_area) %>% 
    mutate(month_year = as.Date(month_year), count = as.numeric(count))
  
  dataFrame[is.na(dataFrame)] <- 0
  
  dataFrame <- dataFrame %>% 
    mutate(type_of_offence = typeOfOffence, count_category = countCategory)
  
  dataFrame
}

plot_data <- function(dataSet, plotTitle) {
  ggplot(dataSet %>% 
  group_by( month_year ) %>% 
  filter( count_category=='Offences' ) %>% 
  summarise( count = sum(count)), aes(x = month_year, y = count)) +
  geom_line(color='red') + ylim(c(0, NA)) + theme_minimal(base_size=10) +
  labs(title=plotTitle,
       subtitle='From January 2009 to September 2016',
       caption='police.govt.nz') +
  xlab('Year') +
  ylab('Number of offences')
}
```

### Offences

Wrangle individual offence sheet data and plot national summary chart for each.

#####Red light running offences at intersections for both vehicles and pedestrians

```{r red_light_officer_issued, echo=FALSE}

#Red light running offences at intersections for both vehicles and pedestrians (seperated by officer issued and red light camera)

red_light_sheet <- read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx", "Red Light")
red_light_select <- red_light_sheet[,select_columns]


offence_officer_red_light <- red_light_select[3:39,]
fines_officer_red_light <- red_light_select[54:90,]

offence_officer_red_light <- restructureData(offence_officer_red_light,"Officer-issued red light","Offences")
fines_officer_red_light <- restructureData(fines_officer_red_light,"Officer-issued red light","Fines")

red_light_offence_and_fines <- rbind(offence_officer_red_light, fines_officer_red_light)

plot_data(red_light_offence_and_fines, "Officer-issued red light offences since 2009")
```


Offences for exceeding the speed limit detected by speed cameras.

```{r speed_camera_issued, echo=FALSE}

# Offences for exceeding the speed limit detected by speed cameras.

speed_camera_sheet <-read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx","Camera-issued Speed")
speed_camera_select <- speed_camera_sheet[,select_columns]

speed_camera_offences <- speed_camera_select[3:39,]
speed_camera_offences <- restructureData(speed_camera_offences,"Camera-issued speed offences","Offences")

speed_camera_fines <- speed_camera_select[45:81,]
speed_camera_fines <- restructureData(speed_camera_fines, "Camera-issued speed offences","Fines")

speed_camera_offences_and_fines <- rbind(speed_camera_offences, speed_camera_fines)

plot_data(speed_camera_offences_and_fines, "Speed camera offences since 2009")


speed_camera_offences %>% filter(month_year >= "2015-10-01") %>% group_by(month_year) %>% 
  summarise(count=sum(count))
```


All restraint use offences (seatbelt and child restraint), excludes restraint equipment offences

```{r restraints, echo=FALSE}

restraints_sheet <- read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx", "Restraints")
restraints_select <- restraints_sheet[,select_columns]

restraints_offences <- restraints_select[3:39,]
restraints_offences <- restructureData(restraints_offences, "Restraints","Offences")

restraints_fines <- restraints_select[45:81,]
restraints_fines <- restructureData(restraints_fines, "Restraints","Fines")

restraints_offences_and_fines <- rbind(restraints_offences, restraints_fines)

plot_data(restraints_offences_and_fines, "Restraint use offences since 2009")
```

```{r alcohol_and_drug, echo=FALSE}

alcohol_and_drugs_sheet <- read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx","Alcohol & Drugs")

alcohol_and_drugs_select <- alcohol_and_drugs_sheet[,select_columns]
alcohol_offences <- alcohol_and_drugs_select[3:39,]
alcohol_offences <- restructureData(alcohol_offences, "Alcohol specific offences","Offences")

drug_offences <- alcohol_and_drugs_select[45:81,]
drug_offences <- restructureData(drug_offences, "Drug specific offences","Offences")

alcohol_fines <- alcohol_and_drugs_select[87:123,]
alcohol_fines <- restructureData(alcohol_fines,"Alcohol specific offences","Fines")

alcohol_and_drugs_offences_and_fines <- rbind(alcohol_offences, drug_offences, alcohol_fines)

plot_data(alcohol_offences,"Alcohol offences since 2009")
```
Offences for using a hand held device for calling or texting while driving


```{r mobile_phone, echo=FALSE}

mobile_phone_sheet <- read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx","Mobile phone")

mobile_phone_select <- mobile_phone_sheet[,select_columns]

mobile_phone_offences <- mobile_phone_select[3:39,]
mobile_phone_offences <- restructureData(mobile_phone_offences, "Mobile phone offences","Offences")

mobile_phone_fines <- mobile_phone_select[45:81,]
mobile_phone_fines <- restructureData(mobile_phone_fines, "Mobile phone offences", "Fines")

mobile_phone_offences_and_fines <- rbind(mobile_phone_offences, mobile_phone_fines)

plot_data(mobile_phone_offences_and_fines,"Mobile phone offences since 2009")

```

```{r all_offences_and_fines}

offences_and_fines <- rbind(alcohol_and_drugs_offences_and_fines, mobile_phone_offences_and_fines, red_light_offence_and_fines, restraints_offences_and_fines, speed_camera_offences_and_fines)

write_csv(offences_and_fines,"../data/offences_and_fines.csv")

write_csv(offences_and_fines %>% 
  group_by(month_year, type_of_offence, count_category) %>% 
  summarise(count =  sum(count)),"../data/national_offences_and_fines.csv")

offences_and_fines$sub_area <- as.factor(offences_and_fines$sub_area)

levels(offences_and_fines$sub_area) 

for(geography in levels(offences_and_fines$sub_area)) {
    temp <- offences_and_fines %>% filter(sub_area == geography)
    filename <- tolower(paste("../../../../workbench/visualisations/police_fines/public/", gsub(" ","_", geography),".csv", sep=""))
    filename <- gsub("-","_",filename)
    write_csv(temp, filename)
}

offences_and_fines %>%
  filter(count_category=='Fines') %>%
  filter(month_year > '2015-08-01') %>%  
  group_by(type_of_offence, count_category) %>%
  summarise(count=sum(count))


offences_and_fines %>%   mutate(year = year(month_year)) %>% 
group_by(year, type_of_offence, count_category) %>%
summarise(count =  sum(count)) %>% ungroup() %>% spread(type_of_offence, count) %>% filter(count_category=="Offences")

```

```{r fleeing_drivers}
fleeing_drivers_sheet <- read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx","Fleeing Drivers")

fleeing_drivers <- fleeing_drivers_sheet[,select_columns]

fleeing_drivers_offences <- fleeing_drivers[3:14,]

fleeing_drivers_offences <- Filter(function(x)!all(is.na(x)), fleeing_drivers_offences)

column_names_fleeing <- c("area",as.character(seq(as.Date("2009/1/1"), as.Date("2016/9/1"), "months")))
names(fleeing_drivers_offences) <- NULL

names(fleeing_drivers_offences) <- column_names_fleeing

fleeing_drivers_offences <- fleeing_drivers_offences %>% gather('date','count', -area) %>% 
   mutate(date = as.Date(date), count = as.numeric(count))

fleeing_drivers_offences[is.na(fleeing_drivers_offences)] <- 0


fleeing_drivers_offences %>% mutate(year = year(date)) %>% group_by(area, year) %>% 
  summarise(count=sum(count))

offences_and_fines_by_year <- offences_and_fines %>%   mutate(year = year(month_year)) %>% 
group_by(year, type_of_offence, count_category) %>%
summarise(count =  sum(count)) %>% ungroup() %>% spread(type_of_offence, count)


write_csv(offences_and_fines_by_year, "../data/offences_and_fines_by_year.csv")


offences_and_fines_by_year %>% filter(count_category=='Offences')

```

```{r fines}
offences_and_fines %>%   mutate(year = year(month_year)) %>%
filter(count_category=="Fines") %>%
group_by(year, type_of_offence, count_category) %>%
summarise(count =  sum(count))  %>% ungroup() %>% group_by(year) %>% summarise(count=sum(count))
```


```{r police_speeding}
police_speeding <- read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx","Police Speeding")

police_speeding_select <- police_speeding[,select_columns]

police_speeding_offences <- police_speeding_select[3:39,]
police_speeding_offences <- restructureData(police_speeding_offences, "Police speeding offences","Offences")

police_speeding_fines <- police_speeding_select[45:81,]
police_speeding_fines <- restructureData(police_speeding_fines, "Mobile phone offences", "Fines")

police_speeding_offences_and_fines <- rbind(police_speeding_offences, police_speeding_fines)

plot_data(police_speeding_offences_and_fines,"Police speeding offences since 2009")
```

