---
title: "Driver offences NZ"
output: html_notebook
---

Driving offence data for NZ - sourced from police.govt.nz

Original file -

The following code reads different sheets from the spreadsheet, combining data for offences and fines by each offence type. There's a main data frame `offences_and_fines` which contains this set.

Note - Sheets are hardcoded, hence tests must be done to double check total values.


```{r load_libraries, include=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library("readxl"))
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library("ggthemes"))
suppressPackageStartupMessages(library("lubridate"))
```


```{r globals, include=FALSE}

# General pattern for offence columns, the gap is to exclude year for consistent renaming. Update the range as new quarterly data comes through
select_columns <- c(1:14,16:27,29:40,42:53,55:66,68:79,81:92,94:102)

column_names <- c("area","sub_area",as.character(seq(as.Date("2009/1/1"), as.Date("2016/9/1"), "months")))


# General function to restructure and reassign frames from different sheets
restructureData <- function(dataFrame, typeOfOffence, countCategory){
  column_names <- c("area","sub_area",as.character(seq(as.Date("2009/1/1"), as.Date("2016/9/1"), "months")))
  
  names(dataFrame) <- column_names
  
  dataFrame <- dataFrame %>% 
    gather('month_year', 'count', -area, -sub_area) %>% 
    mutate(month_year = as.Date(month_year), count = as.numeric(count))
  
  dataFrame[is.na(dataFrame)] <- 0
  
  dataFrame <- dataFrame %>% 
    mutate(type_of_offence = typeOfOffence, count_category = countCategory)
  
  dataFrame
                              
}


plot_data <- function(dataSet, plotTitle) {
  ggplot(dataSet %>% 
  group_by( month_year ) %>% 
  filter( count_category=='Offences' ) %>% 
  summarise( count = sum(count)), aes(x = month_year, y = count)) +
  geom_line(color='red') + ylim(c(0, NA)) + theme_minimal(base_size=10) +
  labs(title=plotTitle,
       subtitle='From January 2009 to September 2016',
       caption='police.govt.nz') +
  xlab('Year') +
  ylab('Number of offences')
  
}
```


### Offences

Wrangle individual offence sheet data and plot national summary chart for each.

#####Red light running offences at intersections for both vehicles and pedestrians

```{r red_light_officer_issued, echo=FALSE}

#Red light running offences at intersections for both vehicles and pedestrians (seperated by officer issued and red light camera)

red_light_sheet <- read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx", "Red Light")
red_light_select <- red_light_sheet[,select_columns]


offence_officer_red_light <- red_light_select[3:39,]
fines_officer_red_light <- red_light_select[54:90,]

offence_officer_red_light <- restructureData(offence_officer_red_light,"Officer-issued red light","Offences")
fines_officer_red_light <- restructureData(fines_officer_red_light,"Officer-issued red light","Fines")

red_light_offence_and_fines <- rbind(offence_officer_red_light, fines_officer_red_light)

plot_data(red_light_offence_and_fines, "Officer-issued red light offences since 2009")
```


Offences for exceeding the speed limit detected by speed cameras.

```{r speed_camera_issued, echo=FALSE}

# Offences for exceeding the speed limit detected by speed cameras.

speed_camera_sheet <-read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx","Camera-issued Speed")
speed_camera_select <- speed_camera_sheet[,select_columns]

speed_camera_offences <- speed_camera_select[3:39,]
speed_camera_offences <- restructureData(speed_camera_offences,"Camera-issued speed offences","Offences")

speed_camera_fines <- speed_camera_select[45:81,]
speed_camera_fines <- restructureData(speed_camera_fines, "Camera-issued speed offences","Fines")

speed_camera_offences_and_fines <- rbind(speed_camera_offences, speed_camera_fines)

plot_data(speed_camera_offences_and_fines, "Camera-issued red light offences since 2009")
```


All restraint use offences (seatbelt and child restraint), excludes restraint equipment offences

```{r restraints, echo=FALSE}

restraints_sheet <- read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx", "Restraints")
restraints_select <- restraints_sheet[,select_columns]

restraints_offences <- restraints_select[3:39,]
restraints_offences <- restructureData(restraints_offences, "Restraints - seatbelts","Offences")

restraints_fines <- restraints_select[45:81,]
restraints_fines <- restructureData(restraints_fines, "Restraints -  seatbelts","Fines")

restraints_offences_and_fines <- rbind(restraints_offences, restraints_fines)

plot_data(restraints_offences_and_fines, "Restraint use offences since 2009")
```

```{r alcohol_and_drug, echo=FALSE}

alcohol_and_drugs_sheet <- read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx","Alcohol & Drugs")

alcohol_and_drugs_select <- alcohol_and_drugs_sheet[,select_columns]
alcohol_offences <- alcohol_and_drugs_select[3:39,]
alcohol_offences <- restructureData(alcohol_offences, "Alcohol specific offences","Offences")

drug_offences <- alcohol_and_drugs_select[45:81,]
drug_offences <- restructureData(drug_offences, "Drug specific offences","Offences")

alcohol_fines <- alcohol_and_drugs_select[87:123,]
alcohol_fines <- restructureData(alcohol_fines,"Alcohol specific offences","Fines")

alcohol_and_drugs_offences_and_fines <- rbind(alcohol_offences, drug_offences, alcohol_fines)

plot_data(alcohol_offences,"Alcohol offences since 2009")
```
Offences for using a hand held device for calling or texting while driving


```{r mobile_phone, echo=FALSE}

mobile_phone_sheet <- read_excel("../data/road-policing-driver-offence-data-jan2009-sep2006.xlsx","Mobile phone")

mobile_phone_select <- mobile_phone_sheet[,select_columns]

mobile_phone_offences <- mobile_phone_select[3:39,]
mobile_phone_offences <- restructureData(mobile_phone_offences, "Mobile phone offences","Offences")

mobile_phone_fines <- mobile_phone_select[45:81,]
mobile_phone_fines <- restructureData(mobile_phone_fines, "Mobile phone offences", "Fines")

mobile_phone_offences_and_fines <- rbind(mobile_phone_offences, mobile_phone_fines)

plot_data(mobile_phone_offences_and_fines,"Mobile phone offences since 2009")
```

